---
title: "Spielberg vs Burton"
author: "Karan Pandanda"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(tidyquant)
```

In this page, we want to explore whether or not the mean IMBD rating for Steven Spielberg is the same as Tim Burton's. We explore this through hypothesis testing and data analysis.


```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)

```

```{r}

Directors_F <- movies %>%
      filter(director =="Steven Spielberg"|director=="Tim Burton")

obs_diff_imdb <- Directors_F %>%
  specify(rating ~ director) %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

null_dist_imdb <- Directors_F %>%
  specify(rating ~ director) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

null_dist_imdb %>% 
  visualize() +
  shade_p_value(obs_stat = obs_diff_imdb, direction = "two-sided")+
  labs(title="Simulation Based Null Distribution")+
  theme_bw()

```
```{r}

null_dist_imdb %>%
  get_p_value(obs_stat = obs_diff_imdb, direction = "two_sided")


```


```{r}

null_dist_imdb  %>%
  get_confidence_interval(point_estimate = obs_diff_imdb, level = .95,type = "se")

```

```{r}

CI_Directors <- Directors_F %>%
  group_by(director) %>%
  summarise(Mean=mean(rating), SE=sd(rating)/sqrt(n()), T.critical = qt(0.975,n()-1),Upper_CI = Mean+SE*T.critical, Lower_CI = Mean-SE*T.critical)


ggplot(CI_Directors, aes(x=Mean, y=reorder(director, Mean)))+
  geom_errorbar(aes(xmin=Lower_CI, xmax=Upper_CI, width=0.2, color=director, size=1.6))+
  labs(title = "Do Spielberg and Burton have the same IMDB ratings?",
       subtitle = "95% confidence intervals overlap",
       x = "Mean IMBD Rating",
       y = "") +
  geom_point(aes(color=director, size=2))+
  geom_rect(aes(xmin=max(Lower_CI), xmax=min(Upper_CI),ymin=-Inf, ymax=+Inf))+
  geom_label(aes(y=reorder(director,Mean),x=Lower_CI,
                label = round(Lower_CI,2)),
                hjust=0.4, vjust=-0.3,
                fill=NA, label.size = NA) +
   geom_label(aes(y=reorder(director,Mean),x=Upper_CI,
                   label = round(Upper_CI,2)),
                hjust=0.4, vjust=-0.3,
                fill=NA, label.size = NA)+
  geom_text(aes(label=round(Mean,2),vjust=-1.5))+
  theme_bw()+
  theme(legend.position = "null")
  

```

Conclusion: Given that the confidence intervals of the 2 means overlap slightly, we could think that the difference between the means is not statistically significant. However, this is misleading as we must look at the p-value in our hypothesis test. The results show a p value of 0.014, and hence, we can reject the null hypothesis that the means are equal. Therefore, the difference between the 2 means is statistically significant and the two directors do not have the same average ratings. 
