---
title: "Covid 19 Mortality"
author: "Karan Pandanda"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(jpeg)
library(png)
library(kableExtra)
library(knitr)  
library(ggplot2)  
library(magick)
```

```{r, include=FALSE}
img <- image_read("/Users/karanpandanda/Downloads/Covid19.jpg")
img <- image_scale(image_scale(img,"500%"),("x100%"))
```

```{r, echo=FALSE}
img
```

## Analysis of mortality rates based on several input variables  

On this page, we are going to look in more detail at the deadliness of Covid-19. We know that the virus has spread across the world over the last few months, sending terror into the minds of billions worldwide. However, is it as deadly as we think it is? Is there evidence to show that people must be extremely worried about this virus. That is what we will be looking at here today; we will observe and analyse data regarding the mortality rate of patients with Covid-19. More specifically, we look to see the impacts of co-morbidities and ICU attendance on the mortality rate across different age categories. We now proceed with our analysis.

We are going to begin by filtering our data-set for the purpose of this analysis: pecifically, we want to see the death rate of patients by gender and sex and whether they had co-morbidities or not, besides whether the patients were admitted to the Intensive Care Unit (ICU) or not.

```{r, cache=TRUE}
# file contains 11 variables and 3.66m rows and is well over 380Mb. 
# It will take time to download

# URL link to CDC to download data
url <- "https://data.cdc.gov/api/views/vbim-akqf/rows.csv?accessType=DOWNLOAD"

covid_data <- read_csv(url)%>% # If vroom::vroom(url) doesn't work, use read_csv(url)
  clean_names()

cleaned_data <- covid_data %>% 
  filter(death_yn %in% c("Yes", "No")) %>% #to remove non-meaningful value in death_yn.
  filter(medcond_yn %in% c("Yes","No") & icu_yn %in% c("Yes", "No")) %>% #to remove non-meaningful value in medcond_yn and icu_yn.
  filter(age_group!="Unknown") %>%  #to remove non-meaningful value in age_group.
  filter(sex %in% c("Male","Female")) %>% 
    mutate(
    medcond_yn=case_when(
      medcond_yn=="Yes" ~ "With comorbidities",
      medcond_yn=="No" ~ "Without comorbidities"
    ),
    icu_yn=case_when(
    icu_yn=="Yes" ~ "Admitted to ICU",
    icu_yn=="No" ~ "No ICU"
    ))
```

Post completion of that, we now start by looking at our first graph. This graph displays the different mortality rates for both men and women, both with and without co-morbidities.

```{r}
cleaned_data %>% 
  group_by(sex, medcond_yn, age_group) %>% 
  summarise(death_num = count(death_yn=="Yes"), total_num =n(), death_rate = death_num/n()) %>% 
  
  ggplot(aes(x=death_rate,y=reorder(age_group, death_rate)))+
    geom_col(fill="#003366")+
    theme_bw() +
    labs (title = "Covid death % by age group, sex and with or without co-morbidities", x="", y="") +
    geom_text(aes(label=round(death_rate*100,1)), hjust=-0.1, size=3)+ #ajust size, position, and decimal for barlabels.
    facet_grid(medcond_yn ~ sex)+ #get the 2x2 combined plot, vertical medcond_yn and horizontal sex.
    scale_x_continuous(limits=c(0,0.85),breaks=seq(0,0.8,0.2),label=scales::percent)+ #make all labels in the plot and  axis label percent.
    theme(plot.margin = unit(c(0.1,0.1,0.1,0.1),"cm"))
```

We can infer some interesting statistics from this graph. There is a clear upward trend in the mortality rate as we go up the age categories, regardless of whether or not the person had a co-morbidity, which was expected as we know elderly people are more susceptible to diseases. However, the increase is exponential, with a near 0 mortality rate for both men and women without co-morbidities till the age of 40, and a shocking 25.2% and 30.7% mortality rate for 80+ men and women without co-morbidities. This indicates that the older you get post 40, the rise in death rates is extremely fast, something policy-makers must take into consideration when looking at designing effective preventative policy. Another thing that can be noted from this graph is that with co-morbidities, the mortality rate is higher for every single age category, regardless of gender. This indicates that this virus also worsens pre-existing conditions, which indicates that people with pre-existing issues should be even more careful. Furthermore, when comparing between genders, we can see that women have a lower death rate in each age category, regardless of the presence of co-morbidities. There are many reasons that could have contributed to this, but primarily, the reason could be that women, on average, tend to be slightly healthier than men. Furthermore, the 0-9 year age category is marginally more susceptible to dying with Covid-19 than the 10-19 category; this is potentially because the immune system hasn't fully developed in the former category, whilst it has in the latter. 

This scenario lends further evidence to the deadliness of this virus, predominantly for the elderly in society. It indicates that policies have to be made to prevent exposure of this 'at-risk' category to this virus and that younger people must ensure to limit interactions with them as well. 

We now want to see whether or not requiring 'Intensive Care Unit' treatment changed the statistics we saw above in any way. We replace the 'co-morbidities' and 'no co-morbidities' with 'ICU' and 'No ICU'. 
```{r}
cleaned_data %>% 
  group_by(sex, icu_yn, age_group) %>% 
  summarise(death_num = count(death_yn=="Yes"), total_num =n(), death_rate = death_num/n()) %>% 
  
  ggplot(aes(x=death_rate,y=reorder(age_group, death_rate)))+
    geom_col(fill="#FF9999")+
    theme_bw() +
    labs (title = "Covid death % by age group, sex and whether patient was admitted to ICU", x="", y="") +
    geom_text(aes(label=round(death_rate*100,1)), hjust=-0.1, size=3)+
    facet_grid(icu_yn ~ sex)+ #get the 2x2 combined plot, vertical icu_yn and horizontal sex.
    scale_x_continuous(limits=c(0,0.85),breaks=seq(0,0.8,0.2),label=scales::percent)+
    theme(plot.margin = unit(c(0.1,0.1,0.1,0.1),"cm"))
```
The results here are unsurprising, but quite horrific to look at. Patients admitted into the ICU clearly had a much higher chance of succumbing to the virus than those who didn't need the ICU. The age-trend we mentioned above continues, with older age categories being more susceptible to the virus. The intuition provided by this graph however, shows just how seriously we should be taking this virus. ICU patients had a mortality rate of 81.1% for males and 75.6% for females (above 80), compared 43.8% and 35.8% of males and females above 80 who didn't need to go to the ICU. Not only does the rate nearly doubles for this age category; it quintiples for the age category 70-79 and is more than 10 times higher for the age category 60-69. Even for the lower age categories, there is a large percentage of deaths that happen in each of them when visiting the ICU, something that needs to be looked at in more detail. This, perhaps, is another reason why there has been an over-crowding of the hospitals and ICUs. The trend of women being less likely to die than men continues here. This graph gives us further intuition behind the deadliness of the virus and how we should prevent its spread. 


## Conclusion

To conclude, we see that mortality rates for patients with co-morbidities and those who have been admitted to the ICU are truly shocking. Although this indicates further evidence behind the deadliness of this virus, we must not discount the mortality rates visible amongst people with no-comorbidities and people who didn't go to the ICU. For males above 80, there is a 43.8% and 30.7% mortality rate respectively for non-ICU admits and those without co-morbidities. This indicates that it isnt just the most vulnerable in society that suffer from this, people without any majorly obvious issues could also succumb to this deadly virus. Hence, we must be careful as a society to limit its spread and protect our elders. 

